---
title: "Programming Assignment 3"
author: "Ana Bennett"
date: "Last updated:`r Sys.Date()`"
output: 
  html_document: 
    highlight: kate
    theme: spacelab

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bivariate Regression: First Attempt

This data set examines the filtered fMRI signal at the most significant voxel and average priming scores for brain-damaged patients, in a study addressing the extent to which phonological and semantic processes recruit the same brain areas.

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(languageR)
``` 

## Tiddying Data
```{r}
imaging %>%
  select(., x = Condition, y = BehavioralScore) 
```


## T-Test

```{r}
t.test(BehavioralScore ~ Condition, data = imaging)
```



## Fitting a Model


```{r}
lm(BehavioralScore ~ FilteredSignal, data = imaging)
  summary(imaging)
```

Many of the basic assumptions regarding a linear model are violated here. For instance, the 1st and 3rd quartiles are not similar; and the mean and median are not near 0. 

This makes sense considering the data. There would be no reason why a filtered voxel signal would explain variance in a behavioral score. Age of brain injury (if it were recorded) would better explain the variance in behavioral score.  

Eventhough this is a significant result, it does not appear to be an important result...


## Plotting with a Regression Line
```{r}
imaging %>%
  select(., FilteredSignal, BehavioralScore)%>%
  ggplot(., aes(x = FilteredSignal, y = BehavioralScore))+
  geom_point()+
  geom_smooth(method = 'lm')
```


## General Observations and Comments

I picked this data set as I am interested in the intersection between language, neuroscience and psychology. However, as I mentioned this data is not captured by a linear model. Considering the variables involved, this makes sense. 

Importantly, this could be avoided by the first step of data analysis -- eyeballing your data with a plot! In addition, the assumptions should be tested if any doubt occurs.  

I will now pick a different data set, first eyeballing the data with a plot to determine that a linear model will best represent this data, redo the steps above. 

##Bivariate Regression: Second Try 

This data set examines durational measurements on the Dutch
prefix ont- in the Spoken Dutch Corpus

##Test that regression model is linear

It appears that the data is 'blob-like' with no apparent patterns
```{r}
library(tidyverse)
library(ggfortify)

mod1 <- lm(SpeechRate ~ DurationPrefixVowel, data = durationsOnt)
summary(mod1)

#regression model is linear
durationsOnt %>%
ggplot(., aes(x = SpeechRate, y = DurationPrefixVowel))+
  geom_point()
```


## Tiddying Data
```{r}
durationsOnt %>%
  select(., x = SpeechRate, y = DurationPrefixVowel)
```


## Correlation

```{r}
cor.test(durationsOnt$DurationPrefixVowel, durationsOnt$SpeechRate, data = durationsOnt)
```

## Plotting with a Regression Line
```{r}
durationsOnt %>%
ggplot(., aes(x = SpeechRate, y = DurationPrefixVowel))+
  geom_point()+
 geom_smooth(method = lm, formula = y ~ poly(x,2)) 
```


## Fitting a Model

``` {r}
lm(DurationPrefixVowel ~ SpeechRate, data = durationsOnt)
summary(durationsOnt)
```


## General Comments

Judging by the plots, and the correlation coefficients, it seems as though there is a small negative relationship between the duration of the vowel and the speech rate. In other words, as speech rate increases, the vowel duration decreases. This makes sense considering the notion that speech influences overall duration of a segment -- thus, a faster speech rate not surprinsingly results in a shorter vowel duration.

However, this result is not significant. I picked vowel duration as I know vowel duration is typically the longer (in comparison to consonants); but perhaps I am omitting a critical variable. For instance, it could be the case that vowel length preceeding plosives is affected, thus a multiple regression would better model this data. 