---
title: "Programming Assignment 4"
subtitle: "HTML Presentation"
author: "Ana Bennett"
date: "`r Sys.Date()`"
output: 
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
       
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r, load libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(ggfortify)
library(devtools)
library(dplyr)
library(knitr)
library(xaringan)
devtools::install_github("yihui/xaringan")
```

Load Data and Tidy
```{r, load data and tidy}
fricatives <- read.csv("fricatives.csv")
fric_tidy <- fricatives %>%
  gather(., key = measurement, val = value, -obs) %>%
  separate(., col = measurement, into = c('phoneme', 'measurement'), sep="_") %>%
 spread(., key = measurement, value = value)
```


---


Descriptive Statistics

```{r, table_print}
fric_tidy %>%
  summarize(., mean_skewness = mean(skewness), sd_skewness = sd(skewness),
mean_cog = mean(cog), sd_cog = sd(cog)) %>% 
  knitr::kable(., format = 'html', digits = 2)
```


---


Center of Gravity as a Function of Phoneme
```{r, boxplot, fig.height = 5.5}
fric_tidy%>%
  select(., cog, phoneme)%>%
  ggplot(., aes(x = phoneme, y = cog, fill = phoneme))+
  geom_boxplot()+
  labs(x = "phoneme", y = "center of gravity", 
       title = "Center of Gravity as a Function of Phoneme")
```


----


Skewness as a Function of Phoneme
```{r, transformation, fig.height= 4}
fric_tidy%>%
  select(.,skewness, phoneme, cog)%>%
  ggplot(., aes(x = phoneme, y = skewness, color = cog)) +
  geom_jitter()+
  labs( x = "phoneme", y = "skewness", title = "Skewness as a Function of Phoneme")
```


---


Center of Gravity as a Function of Skewness: the phoneme [s]


The data
```{r, new data set for [s]}
fricatives <- read.csv("fricatives.csv")
s_data <- fricatives %>%
  gather(., key = measurement, val = value, -obs) %>%
  separate(., col = measurement, into = c('phoneme', 'measurement'), sep="_") %>%
  spread(., key = measurement, value = value) %>%
  select(., cog, skewness, phoneme)%>%
  filter(., phoneme == 's')

head(s_data)
```


---


Model summary for Center of Gravity and Skewness: the phoneme [s]
```{r, modeling, message=FALSE, warning=FALSE}
my_mod <- lm(cog ~ skewness, data = s_data)
summary(my_mod)
mean(my_mod$residuals)
```


---


Scatterplot: Center of Gravity and Skewness
```{r, scatter, message=FALSE, warning=FALSE, fig.height= 5}
s_data %>%
  select(., cog, skewness)%>%
  ggplot(., aes(x = cog, y = skewness))+
  geom_point(method = 'lm')+
  geom_smooth()
  labs(x = "Center of Gravity", y = "Skewness", title = 
         "Center of Gravity as a Function of Skewness")
```


---


Model diagnositics: Overview of Normality 


```{r, message=FALSE, warning=FALSE}
autoplot(my_mod)
```


---


Durbin-Watson Test


```{r, message=FALSE, warning=FALSE}
lmtest::dwtest(my_mod)
```


---


Normality test of Residuals: QQplot


```{r, message=FALSE, warning=FALSE}
autoplot(my_mod, which = 2)
```


---


Check for autocorrelation of residuals


```{r, message=FALSE, warning=FALSE}
acf(my_mod$residuals)
```


---


Interpretation of Results

Statistics Overview

A linear model was used to fit data regarding fricatives( [s] and [sh]), center of gravity, and skewness. To check for normality, the following tests were
implmented: a homoskedasticity test of residuals, a check for autocorrelation of residuals, a normality plot for residuals, and a Durbin-Watson test. Overall,
these tests indicated normality of the data. 

General Conclusions
To better understand the data three plots were made: (1) one with center of gravity as a function of phoneme; (2) and another with skewness as a function of phoneme, (3) a scatterplot testing corelation between skewness and center of gravity.Plot (1) indicated that cog tends to have a higher value for [s] in comparison to [sh]. While plot (2)indicated that the phoneme is related to the direction of skewness. In other words, with the fricative [s] there is a negative skew; and with [sh] we see a positive skew. Lastly, in plot (3) there was a positive correlation between skewness and center of gravity. Futhermore, when we tested the model for cog and skewness (specifcally, for the phoneme [s]), we found that skewness was significant (p <0.001). This suggests that skewness is a good predictor for cog. 








